Honors Thesis

### 2009-02-27 honors_thesis/paper.tex
Working on the paper

@@

\section{Methodology/Procedures}

@@

@@

@@

>>>>

Optimization problems are ubiquitous in our modern world.  Businesses such as
Google need to decide where to place ads, scientists need to fit models to
data, and airlines need to schedule their flights.  All of those problems are
optimization problems, and optimization techniques can be used to find
solutions.  Optimization algorithms are derived from all sorts of natural
phenomena.  Some methods look at the landscape of the function as they are
searching and follow hills to the best values, as in the gradient ascent
method.  Other methods take ideas from biology or metallurgy, like genetic
algorithms and simulated annealing.  Particle swarm optimization is a recently
developed optimization technique that draws on ideas from the sociology of
flocking birds.  All of these methods, however, are fundamentally sequential in
nature---they need to be run in a specific order, and because of that, they are
typically only run on one machine.  The problems people want to solve are
getting bigger, and larger problems need to run on multiple machines in
parallel.  Parallel implementations of these algorithms have been attempted,
but they do not often perform as well as sequential versions for the amount of
computation performed.  The work I propose focuses on one of those algorithms,
particle swarm optimization, and improving its performance in a large-scale,
parallel world.

@@

I will study the feasibility and potential performance gain of speculative
execution in particle swarm optimization.  By speculative execution I mean the
use of additional computational resources, presumably extra processors running
in parallel, to perform two or more iterations of particle swarm optimization
at a time.

\section{Methodology}

The methodology of this research will procede as follows:

Compare to original How to control for external factors (nice in CS, though not
in other sciences) Once you've run experiments, what are you looking for?
Anything you can say about what code you're using, and how you intend to
compare results.  People are concerned about how you're going to draw
conclusions.  It matters what kind of statistical tests you are going to do.
We're going to run a number of trials because it's a randomized algorithm.
Test against multiple functions, with multiple topologies - some synthetic
functions, and some real-world applications.  Run with standard constriction
coefficient.  How do you compare your results to things other people have done?

\begin{itemize}

\item

\end{itemize}

@@

I am a Computer Science major.  My main focuses in my undergraduate career have
been optimization, natural language processing, and machine learning.  The
optimization emphasis has come mainly in my work with Dr. Kevin Seppi, for whom
I am a research assistant.  I have spent the last year working with him,
working mainly on particle swarm optimization topics.  We worked together with
Dr. Branton Campbell, a physics professor, to solve a crystallography problem
for him using a parallel implementation of PSO.  Those results will be
submitted to Nature Materials.  Recently we, along with a graduate student who
works in the lab, submitted a paper to the Congress on Evolutionary
Computation, a major venue for optimization techniques.  The paper was on
topologies and communication in large particle swarms, and formed the
beginnings of the work I propose.

@@

Dr. Kevin Seppi is the head of the Applied Machine Learning Laboratory at BYU.
Three of his last five graduate students have done their graduate work in
particle swarm optimization, and he has authored or co-authored over 10 papers
on the topic.  I took CS 236 from him in the fall of 2007 and decided I wanted
to do research in the area he was researching, so I joined his lab.  I too
became interested in PSO and have worked in that area for the past year.

@@

This project will not require any outside funding.  The only expenditure
will be my own time researching, conducting experiments, and writing.
